\textit{MSF: Modulated Sub-graph Finder}

### 2018-03-20 MSF_Paper_FA.tex
a few paragraphs more

On these grounds we propose a novel approach to make use of the rich
pathway annotation resources available to gain additional functional
insights from basic DGEA. To this, we start with the presupposition
that expression of neighboring genes within a functional pathway are
not independent from each other. Rather, they are often regulating
each others expression or are part of the same
regulon~\cite{Michalak}. We understand that the categorization of
links between genes into labeled pathways is often an arbitrary one,
given the extensive cross talk between different pathways. Although
these categories have proven to be useful in many situations, they
force a certain perspective onto the interpretation of novel
data. Based on these two principles, we aim to find sub-graphs of
connected genes within cell signaling network which exhibit as a whole
significant differential expression changes. As input, information on
functional links between genes provided by e.g. KEGG or Reactome and
information on the differential expression status of single genes
resulting from a DGEA, are required. As a result the analysis returns
sub-graphs and their joint confidence scores, reflecting how the
perturbation is migrated through the network. Furthermore, the entry
points of perturbation in the networks and overlap with conventional
pathway categories are returned. All of this can be helpful to
understand the cause end effect of a stimulus and might inform about
potential points of intervention. The proposed algorithm was
implemented as a java program, which was named Modulated Sub-graph
Finder (\texttt{MSF}). \texttt{MSF} would help transform the
information obtained from DGEA into understandable knowledge of signal
transduction of genes. \texttt{MSF} is freely accessible on github.

>>>>

On these grounds we propose a novel approach to make use of the rich
pathway annotation resources available to gain additional functional
insights from basic DGEA. To this, we start with the presupposition
that expression of neighboring genes within a functional pathway are
not independent from each other. Rather, they are often regulating
each others expression or are part of the same
regulon~\cite{Michalak}. We understand that the categorization of
links between genes into labeled pathways is often an arbitrary one,
given the extensive cross talk between different pathways. Although
these categories have proven to be useful in many situations, they
force a certain perspective onto the interpretation of novel
data. Based on these two principles, we aim to find sub-graphs of
connected genes within cell signaling network which exhibit as a whole
significant differential expression changes. As input, information on
functional links between genes provided by e.g. KEGG or Reactome and
information on the differential expression status of single genes
resulting from a DGEA, are required. As a result the analysis returns
sub-graphs and their joint confidence scores, reflecting how the
perturbation is migrated through the network. Furthermore, the entry
points of perturbation in the networks and overlap with conventional
pathway categories are returned. All of this can be helpful to
understand the cause and effect of a stimulus and might inform about
potential points of intervention. The proposed algorithm was
implemented as a java program, which was named \texttt{Modulated Sub-graph
Finder} (abbreviated \texttt{MSF}). \texttt{MSF} can help transform the
information obtained from DGEA into understandable knowledge of signal
transduction of genes and thereby being a valuable complement to existing
pathway based methods. \texttt{MSF} is freely accessible on github.
